Ultrafast Electron Microscopes: Design Criteria, Electron Sources, and Column Modeling

### 2013-04-11 inc/gun/gun_model.tex
added cc notices to the tops of several files
more corrections and punctuation
citations are now preceeded by non-breaking space (~)

To model an acceleration region, one must assume a form for the accelerating field.
From that form one may derive the terms governing the pulse dynamics. Initially, assume that the on-axis field is purely in the longitudinal direction, which may be thought of as an implication of radial symmetry.
Mathematically, that is
\begin{equation}
  \vec{F}(r^{\prime}=0,z^{\prime}) = F_z (0,z^{\prime}) \hat{z} = e E_z (0,z^{\prime}) \hat{z}
\end{equation}
where as before, the primed coordinates are the location of the center of the pulse.

To first model the longitudinal (compression) action at the anode, %TODO ref
consider electrons that are a distance $z$ ahead or behind of the center of the pulse and so feel a slightly different force, 
\begin{equation}
  F_z(0,z^{\prime} - z) \approx F_z(0,z^{\prime} - z) + \frac{\partial F_z}{\partial z^{\prime}} z
\end{equation}
Again, the constant leading term does not contribute to the intrapulse dynamic, however the second term is of the proper form to say that in this region the anode's longitudinal compression strength $M_z$ is given by
\begin{equation}
  M_z = \frac{\partial F_z}{\partial z^{\prime}} = e \frac{\partial E_z}{\partial z^{\prime}} \, \text{.}
\end{equation}

To model the radial action (negative transverse lensing), we can expand the radial force $F_r$ for electrons that are a distance $r$ radially outward from the center of the pulse,
\begin{equation}
  F_r(r,z^{\prime}) \approx F_r(r,z^{\prime}) + \frac{\partial F_r}{\partial r^{\prime}} r
\end{equation}
Without an analytic form for $F$, we can appeal to Gauss' law (in absense of external charge)
\begin{gather}
  \frac{\partial F_x}{\partial x^{\prime}} + \frac{\partial F_y}{\partial y^{\prime}} + \frac{\partial F_z}{\partial z^{\prime}} = 2 \frac{\partial F_{\smallT}}{\partial r^{\prime}} + \frac{\partial F_z}{\partial z^{\prime}} = 0 \\
  \Rightarrow \frac{\partial F_{\smallT}}{\partial r^{\prime}} = - \frac{1}{2} \frac{\partial F_z}{\partial z^{\prime}}
\end{gather} 
to see that the transverse lens traverse lens strength $M_{\smallT}$ from the anode is simply,
\begin{equation}
  M_{\smallT} = -\frac{1}{2}\frac{\partial F_z}{\partial z^{\prime}} \, \text{.}
\end{equation}

We may choose to assume that the on-axis field is defined as a smoothed step down function
\begin{equation} \label{eq:anode_tanh}
  F_z(0,z^{\prime}) = \frac{eV}{2z^{\prime}_A} \left( 1 - \tanh \left( \frac{ z^{\prime} - z^{\prime}_A }{ z^{\prime}_A / s } \right) \right)
\end{equation}
where $V$ is the accelerating potential, $z^{\prime}_A$ is the position of the anode (the cathode assumed to be at $z^{\prime} = 0$) and $s$ is a constant quantifying the ``sharpness'' of the fall-off of the electric field.
Indeed, this form is analgous to the super-Gaussian envalope (\ref{eq:reg_of_influence}), however it is open ended on the cathode side, which is necessary for the electrons to be accelerated away from the photocathode.
Given this form the ``lensing'' strengths are then
\begin{gather}
  M_{\smallT} = -\frac{eVs}{4z^{\prime 2}_A} \operatorname{sech}^2 \left( \frac{z^{\prime} - z^{\prime}_A }{ z^{\prime}_A / s } \right) \\
  M_z = \frac{eVs}{2z^{\prime 2}_A} \operatorname{sech}^2 \left( \frac{z^{\prime} - z^{\prime}_A }{ z^{\prime}_A / s } \right) \, \text{.}
\end{gather}
As with the RF cavity, one needs to include the on-axis force (\ref{eq:anode_tanh}) in the overall pulse propagation kinematics.

@@

\ref{fig:gun-sharpness} shows an example acceleration region for three values of the ``sharpness'' $s$.
What is clear from this figure is that the dynamics are not highly coupled to the value of this hard-to-quantify parameter.
For all values of $s$ the divergence of the beam is equal (though the size may vary slightly).
Interestingly the length of the pulse is stabilized on exiting the anode, coming very nearly to a constant length (not expanding nor contracting).
This effect is due to the slower electrons remaining in the cavity longer and thus receiving an additional kick, as described in \cite{oudheusden_electron_2007}.
This longitudinal electric field shape is compared to the electric field computed by finite-element (FEM) simulations of our acceleration cavity in Section \ref{sec:gun_design}.

>>>>

To model an acceleration region, one must assume a form for the accelerating field.
From that form one may derive the terms governing the pulse dynamics. Initially, assume that the on-axis field is purely in the longitudinal direction, which may be thought of as an implication of radial symmetry.
Mathematically, that is
\begin{equation}
  \vec{F}(r^{\prime}=0,z^{\prime}) = F_z (0,z^{\prime}) \hat{z} = e E_z (0,z^{\prime}) \hat{z} \,\text{,}
\end{equation}
where as before, the primed coordinates are the location of the center of the pulse.

To first model the longitudinal (compression) action at the anode, %TODO ref
consider electrons that are a distance $z$ ahead or behind of the center of the pulse and so feel a slightly different force, 
\begin{equation}
  F_z(0,z^{\prime} - z) \approx F_z(0,z^{\prime} - z) + \frac{\partial F_z}{\partial z^{\prime}} z \,\text{,}
\end{equation}
Again, the constant leading term does not contribute to the intrapulse dynamic, however the second term is of the proper form to say that in this region the anode's longitudinal compression strength $M_z$ is given by
\begin{equation}
  M_z = \frac{\partial F_z}{\partial z^{\prime}} = e \frac{\partial E_z}{\partial z^{\prime}} \, \text{.}
\end{equation}

To model the radial action (negative transverse lensing), we can expand the radial force $F_r$ for electrons that are a distance $r$ radially outward from the center of the pulse,
\begin{equation}
  F_r(r,z^{\prime}) \approx F_r(r,z^{\prime}) + \frac{\partial F_r}{\partial r^{\prime}} r \,\text{.}
\end{equation}
Without an analytic form for $F$, we can appeal to Gauss' law (in absense of external charge),
\begin{gather}
  \frac{\partial F_x}{\partial x^{\prime}} + \frac{\partial F_y}{\partial y^{\prime}} + \frac{\partial F_z}{\partial z^{\prime}} = 2 \frac{\partial F_{\smallT}}{\partial r^{\prime}} + \frac{\partial F_z}{\partial z^{\prime}} = 0 \\
  \Rightarrow \frac{\partial F_{\smallT}}{\partial r^{\prime}} = - \frac{1}{2} \frac{\partial F_z}{\partial z^{\prime}} \,\text{,}
\end{gather} 
to see that the transverse lens traverse lens strength $M_{\smallT}$ from the anode is simply,
\begin{equation}
  M_{\smallT} = -\frac{1}{2}\frac{\partial F_z}{\partial z^{\prime}} \, \text{.}
\end{equation}

We may choose to assume that the on-axis field is defined as a smoothed step down function
\begin{equation} \label{eq:anode_tanh}
  F_z(0,z^{\prime}) = \frac{eV}{2z^{\prime}_A} \left( 1 - \tanh \left( \frac{ z^{\prime} - z^{\prime}_A }{ z^{\prime}_A / s } \right) \right) \,\text{,}
\end{equation}
where $V$ is the accelerating potential, $z^{\prime}_A$ is the position of the anode (the cathode assumed to be at $z^{\prime} = 0$) and $s$ is a constant quantifying the ``sharpness'' of the fall-off of the electric field.
Indeed, this form is analgous to the super-Gaussian envalope (\ref{eq:reg_of_influence}), however it is open ended on the cathode side, which is necessary for the electrons to be accelerated away from the photocathode.
Given this form the ``lensing'' strengths are then
\begin{gather}
  M_{\smallT} = -\frac{eVs}{4z^{\prime 2}_A} \operatorname{sech}^2 \left( \frac{z^{\prime} - z^{\prime}_A }{ z^{\prime}_A / s } \right) \\
  M_z = \frac{eVs}{2z^{\prime 2}_A} \operatorname{sech}^2 \left( \frac{z^{\prime} - z^{\prime}_A }{ z^{\prime}_A / s } \right) \, \text{.}
\end{gather}
As with the RF cavity, one needs to include the on-axis force (\ref{eq:anode_tanh}) in the overall pulse propagation kinematics.

@@

\ref{fig:gun-sharpness} shows an example acceleration region for three values of the ``sharpness'' $s$.
What is clear from this figure is that the dynamics are not highly coupled to the value of this hard-to-quantify parameter.
For all values of $s$ the divergence of the beam is equal (though the size may vary slightly).
Interestingly the length of the pulse is stabilized on exiting the anode, coming very nearly to a constant length (not expanding nor contracting).
This effect is due to the slower electrons remaining in the cavity longer and thus receiving an additional kick, as described in~\cite{oudheusden_electron_2007}.
This longitudinal electric field shape is compared to the electric field computed by finite-element (FEM) simulations of our acceleration cavity in Section \ref{sec:gun_design}.
